`State and Parameter Estimation for Spatio-Temporal Neural Fields' or `A Data Driven Framework for Patient-Specific Neural Field Modelling'

### 2010-06-21 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
minor edits, commiting for merge

\subsection{Derivation of Finite Dimensional State-Space Model} 

@@

   	\caption{An example of continuous field decompposed by a finite number of basis functions scaled by the state vector. } 

@@

\section{Discussion}\label{DiscussionSection}
This paper has presented a method for estimating intracortical connectivity and synaptic dynamics from electrophysiological data. The results indicate good estimation performance, where the mean parameter estimates are in good accordance to the actual values. The estimates of the connectivity kernel parameters show relatively broad distributions, which is expected considering the system is driven by an unknown input, modelled as a random disturbance and the observations are corrupted by noise. To overcome the effect of noise, an evoked-response paradigm should be considered when using electrophysiological recordings. Following such a paradigm, data can be epoched into a sufficient number of realisations, where each realisation is a response to a stimuli, such that a distribution of the parameter estimates can be formed.

A significant contribution of this paper is the derivation of the basis function decomposition of the neural field model. Through the derivation, a finite-dimensional state-space model is formed that describes a continuous neural field. Previous work from Schiff and Sauer~\ref{schiff2008kalman} demonstrated for the first time the applicability of the unscented Kalman filter for state-estimation of a discrete (in space and time) nonlinear spatial-temporal systems. Within their framework, each discrete spatial location forms an element of the state vector to be estimated with the UKF. Therefore, a limitation of this framework lies in an increase in dimensionality with increases in  the spatial resolution and in the physical size of the neural field. Increases in dimensionality lead to increases in the uncertainty the estimated state, and problems associated with computational complexity. This paper builds on this previous study, where the estimator is constructed to describe a continuous field. The dimensionality of the state vector is independent of the resolution of a spatial discretisation and is governed by the inherent spatial properties of the field. The frequency analysis methods provide a rigorous framework for the placement and the number of the basis functions required for the decomposition. This enables estimation of a larger field, with a lower dimensional state vector. The basis function decomposition also enabled the coupling to be described by an inner product rather than a spatial convolution. This transformation greatly simplifies the computational complexity of the estimator, which is important for the estimator to run in a real-time scenario.

The frequency analysis and model selection tools presented in this paper specify the sensor arrangement that is required for capturing the dominant cortical dynamics. Importantly, these new tools can also be used to better design electrophysiological recording techniques to avoid spatial aliasing. This analysis is becoming increasingly relevant as recording systems are becoming more sophisticated with higher density recordings~\cite{Brinkmann2009}.

To the authors' best knowledge, this is the first paper to propose an estimation framework for intracortical connectivity from electrophysiological data. The importance of the estimation framework can be illustrated by considering that the connectivity structure at this scale can not be estimated using diffusion tensor imaging due to the near isotropic diffusion of water in grey matter~\cite{Assaf2008}.

>>>>

\subsection{Derivation of Finite Dimensional State-Space Model}\section{Sect:ReducedModelDerivation}

@@

   	\caption{An example of continuous field decomposed by a finite number of basis functions scaled by the state vector. } 

@@

\section{Discussion}\label{DiscussionSection}
This paper has presented a method for estimating intracortical connectivity and synaptic dynamics from electrophysiological data. The results indicate good estimation performance, where the mean parameter estimates are in good agreement to the actual values. The estimates of the connectivity kernel parameters show relatively broad distributions, which is expected considering the system is driven by an unknown input, modelled as a random disturbance and the observations are corrupted by noise. To overcome the effect of noise, an evoked-response paradigm should be considered when using electrophysiological recordings. Following such a paradigm, data can be epoched into a sufficient number of realisations, where each realisation is a response to a stimuli, such that a distribution of the parameter estimates can be formed.

A significant contribution of this paper is the derivation of the basis function decomposition of the neural field model, as described in Section~\ref{Sect:ReducedModelDerivation}. Through the derivation, a finite-dimensional state-space model is formed that describes a continuous neural field. Previous work from Schiff and Sauer~\ref{schiff2008kalman} demonstrated for the first time the applicability of the unscented Kalman filter for state-estimation of a discrete (in space and time) nonlinear spatial-temporal systems. Within their framework, each discrete spatial location forms an element of the state vector to be estimated with the UKF. Therefore, a limitation of this framework lies in an increase in dimensionality with increases in  the spatial resolution and in the physical size of the neural field. Increases in dimensionality lead to increases in the uncertainty the estimated state, and problems associated with computational complexity. This paper builds on this previous study, where the estimator is constructed to describe a continuous field. The dimensionality of the state vector is independent of the resolution of a spatial discretisation and is governed by the inherent spatial properties of the field. The frequency analysis methods provide a rigorous framework for the placement and the number of the basis functions required for the decomposition. This enables estimation of a larger field, with a lower dimensional state vector. The basis function decomposition also enabled the coupling to be described by an inner product rather than a spatial convolution. This transformation greatly simplifies the computational complexity of the estimator, which is important for the estimator to run in a real-time scenario.

The frequency analysis and model selection tools presented in Section~\ref{SpectralAnalysisSection} specifies the sensor arrangement that is required for capturing the dominant cortical dynamics. Importantly, these new tools can also be used to better design electrophysiological recording techniques to avoid spatial aliasing. This analysis is becoming increasingly relevant as recording systems are becoming more sophisticated with higher density recordings~\cite{Brinkmann2009}.

To the authors' best knowledge, this is the first paper to propose a model-based estimation framework for intracortical connectivity from electrophysiological data. The importance of the estimation framework can be illustrated by considering that the connectivity structure at this scale can not be estimated using diffusion tensor imaging due to the near isotropic diffusion of water in grey matter~\cite{Assaf2008}.
